What is the impact of weather catastrophe frequency and severity on property insurance policy prices, and how can homeowners alleviate financial strain?

### 2023-10-30 manuscript/manuscript.tex
tables and equations added to data section
added method

\section{Data}
\label{sec:data}
To comprehend the impact of weather-related catastrophic events on property insurance pricing, we will begin by examining a dataset of 
billion-dollar disasters. This dataset provides crucial information on event frequency, financial cost, and other key parameters. Our 
data sources include the National Centers for Environmental Information (NCEI)\cite{ncei}, the Insurance Information Institute 
(III)\cite{iii}, Aon\cite{aon} and the National Association of Insurance Commissioners (NAIC)\cite{naic}, all are recognized for their 
reliability in documenting catastrophic events and their associated economic and insured losses.

@@

\begin{table}[h]
  \label{tab:natural_cat_losses}
    \centering
    \caption{Natural Catastrophe Losses in the United States by Peril, 2022 (in \textdollar{} millions)}
    \begin{tabular}{|l|c|c|c|c|}
    \hline
    Peril & Number of Events & Fatalities & Economic Losses (2) & Insured Losses (3) \\
    \hline
    Tropical Cyclone & 3 & 157 & \textdollar 96,097 & \textdollar 53,203 \\
    Severe Convective Storm & 62 & 49 & \textdollar 37,232 & \textdollar 29,306 \\
    Wildfire, Drought, Heatwave & 26 & 65 & \textdollar 18,093 & \textdollar 8,902 \\
    Winter Storm & 13 & 123 & \textdollar 6,223 & \textdollar 4,128 \\
    Flooding & 15 & 72 & \textdollar 7,234 & \textdollar 3,346 \\
    Total & 119 & $\sim$466 & \textdollar 164,879 & \textdollar 98,885 \\
    \hline
    \end{tabular}    
    \cite{iii}
\end{table}

@@

\begin{table}[h]
  \label{tab:insured_prop_losses}
    \centering
    \caption{Estimated Insured Property Losses, U.S. Natural Catastrophes, 2013-2022 (in \textdollar billions)}
    \begin{tabular}{|c|c|c|}
    \hline
    Year & In dollars when occurred & In 2022 dollars (2) \\
    \hline
    2013 & \textdollar 24.1 & \textdollar 31.0 \\
    2014 & \textdollar 23.2 & \textdollar 29.2 \\
    2015 & \textdollar 22.9 & \textdollar 28.8 \\
    2016 & \textdollar 31.6 & \textdollar 39.3 \\
    2017 & \textdollar 130.9 & \textdollar 158.7 \\
    2018 & \textdollar 60.4 & \textdollar 71.6 \\
    2019 & \textdollar 38.8 & \textdollar 45.2 \\
    2020 & \textdollar 81.0 & \textdollar 93.3 \\
    2021 & \textdollar 93.3 & \textdollar 102.7 \\
    2022 & \textdollar 98.8 & \textdollar 99.9 \\
    \hline
    \end{tabular}
    \cite{iii}
\end{table}

@@

\section{Methods}
\label{sec:meth}

>>>>

\section{Data}
\label{sec:data}
To comprehend the impact of weather-related catastrophic events on property insurance pricing, we will begin by examining a dataset of 
billion-dollar disasters. This dataset provides crucial information on event frequency, financial cost, and other key parameters. Our 
data sources include the National Centers for Environmental Information (NCEI)\cite{ncei}, the Insurance Information Institute 
(III)\cite{iii}, Aon\cite{aon} and the National Association of Insurance Commissioners (NAIC)\cite{naic21}, all are recognized for their 
reliability in documenting catastrophic events and their associated economic and insured losses.

@@

\begin{table}[h]
  \label{tab:natural_cat_losses}
    \centering
    \begin{tabular}{|l|c|c|c|c|}
        \hline
        Peril & Number of Events & Fatalities & Economic Losses (2) & Insured Losses (3) \\
        \hline
        Tropical Cyclone & 3 & 157 & \textdollar 96,097 & \textdollar 53,203 \\
        Severe Convective Storm & 62 & 49 & \textdollar 37,232 & \textdollar 29,306 \\
        Wildfire, Drought, Heatwave & 26 & 65 & \textdollar 18,093 & \textdollar 8,902 \\
        Winter Storm & 13 & 123 & \textdollar 6,223 & \textdollar 4,128 \\
        Flooding & 15 & 72 & \textdollar 7,234 & \textdollar 3,346 \\
        Total & 119 & $\sim$466 & \textdollar 164,879 & \textdollar 98,885 \\
        \hline
    \end{tabular}    
    \caption{Natural Catastrophe Losses in the United States by Peril, 2022 (in \textdollar millions)}
    \cite{iii}
\end{table}

@@

\begin{table}[h]
  \label{tab:insured_prop_losses}
    \centering
    \begin{tabular}{|c|c|c|}
        \hline
        Year & In dollars when occurred & In 2022 dollars (2) \\
        \hline
        2013 & \textdollar 24.1 & \textdollar 31.0 \\
        2014 & \textdollar 23.2 & \textdollar 29.2 \\
        2015 & \textdollar 22.9 & \textdollar 28.8 \\
        2016 & \textdollar 31.6 & \textdollar 39.3 \\
        2017 & \textdollar 130.9 & \textdollar 158.7 \\
        2018 & \textdollar 60.4 & \textdollar 71.6 \\
        2019 & \textdollar 38.8 & \textdollar 45.2 \\
        2020 & \textdollar 81.0 & \textdollar 93.3 \\
        2021 & \textdollar 93.3 & \textdollar 102.7 \\
        2022 & \textdollar 98.8 & \textdollar 99.9 \\
        \hline
    \end{tabular}
    \caption{Estimated Insured Property Losses, U.S. Natural Catastrophes, 2013-2022 (in \textdollar billions)}
    \cite{iii}
\end{table}

@@

\section{Methods}
\label{sec:meth}
We estimate the influence of natural disasters on homeowners insurance premiums using panel regression models with state and year 
fixed effects:

\begin{equation} 
    \mathrm{Premium}{ist} = \beta_0 + \beta_1 , \mathrm{Disasters}{ist} + \gamma_i + \delta_t + \epsilon_{ist} 
\end{equation}

The dependent variable is the logged average premium in state $i$ and year $t$. The main independent variables are disaster measures 
for state-year:

\begin{itemize} 
    \item Number of events 
    \item Total cost 
    \item Insured losses 
    \item Indicators for peril types: hurricane, flood, severe storm, winter weather, wildfire, earthquake 
\end{itemize}

Logs are used given the skewed distribution of disaster cost variables. State and year fixed effects control for time-invariant 
differences across states and national trends. We first examine the influence of aggregated disaster activity. We then estimate 
models interacting the disaster measures with peril indicators to compare sensitivity across event types. Quantile regression is 
also used to test if effects differ across the premium distribution.
